Project Proposal

### 2024-11-17 CURJ_IDS702Projectproposal.qmd
commit change to branch

**Brief description:** The source of the datasets includes health
factor and health outcome data for each county in the US by state in
2024. These datasets are provided as part of a program by the University
of Wisconsin Population Health Institute and combines data from other
sources, such as the Behavioral Risk Factor Surveillance System &
National Center for Health Statistics. We are compiling the datasets for
seven states in seven different regions of the US to create one dataset
(North Carolina, Texas, Alabama, Colorado, Minnesota, Washington, & New
York). We will also be merging in a second American Community Survey dataset to obtain property tax data for our second research question.

**Research question 1:** What is the relationship between residential
segregation (continuous variable), percentage of the county population
with limited access to healthy foods (continuous variable), and
percentage of adults who are uninsured (continuous variable), and
average life expectancy (continuous variable)? How is the relationship
between residential segregation and life expectancy affected by the
region of the US that the county / state is in (categorical variable
that we will create called Region)? The outcome variable is average life
expectancy. The predictors are residential segregation, percentage with
limited access to healthy foods, and percentage of adults who are
uninsured.

Average life expectancy = average number of years a person is expected
to live (data from National Center for Health Statistics - Natality and
Mortality Files; Census Population Estimates Program, 2019-2021).

Residential segregation = Index of dissimilarity where higher values
indicate greater residential segregation between Black and white county
residents (data from American Community Survey 5-year estimate,
2018-2022).

Percentage limited access to healthy foods = Percentage of population
who are low-income and do not live close to a grocery store (data from
USDA Food Environment Atlas, 2019).

Percentage adults uninsured = Percentage of adults under age 65 without
health insurance (data from Small Area Health Insurance Estimates,
2021).

@@

The outcome variable is school funding adequacy, with 0=inadequate
funding and 1=adequate funding. The predictors are residential
segregation, property tax, and median household income.

School funding adequacy = The average gap in dollars between actual and
required spending per pupil among public school districts. Required
spending is an estimate of dollars needed to achieve U.S. average test
scores in each district. (data from School Finance Indicators Database,
2021).

Median property tax = Median property tax value for each county between
2018-2022 (data from IPUMS).

Median household income = The income where half of households in a
county earn more and half of households earn less (data from Small Area
Income and Poverty Estimates; American Community Survey, 5-year
estimates, 2022 & 2018-2022).

**Load the datasets and prepare final dataset`**:

@@

**Provide a `glimpse():**

>>>>

**Brief description:** This dataset contains health factor and health
outcome data for each county in NC in 2024. It is provided as part of a
program by the University of Wisconsin Population Health Institute ()
and combines data from other sources, such as the Behavioral Risk Factor
Surveillance System & National Center for Health Statistics.

**Research question 1:** What is the relationship between residential
segregation (continuous variable) and juvenile arrests (continuous
variable)? How is this relationship affected by the region of the US
that the county / state is in (categorical variable called Region)?

-   **Outcome (Dependent) Variable** $Juvenile.Arrest.Rate$

    -   Represents the rate of delinquency cases per 1,000 juveniles in
        a county. This variable quantifies juvenile arrests as a measure
        of youth criminal activity.
    -   **Source**: Data from Juvenile Justice Databases, 2021.

-   **Primary Independent Variable** $Residential.Segregation.Index$

    -   The index of dissimilarity measures residential segregation
        between Black and white residents in a county. Higher values
        indicate greater segregation. This metric captures the extent to
        which these populations reside in separate geographic areas
        within the same county.
    -   **Source:** American Community Survey 5-year estimates,
        2018-2022.

-   **Interaction Term** $Residential.Segregation.Index * Region$

    -   Examines how the relationship between residential segregation
        and juvenile arrests varies by region (`Region` categorical
        variable).

-   **Control Variables** Adjusts for other factors that might confound
    the relationship between segregation and arrests.

    -   $Spending.per.Pupil$ –The average educational expenditure per
        student in a district, which reflects investment in education
        and resources that may impact juvenile arrest rates.

        -   **Source:** State or National Education Finance Databases,
            2021

    -   $School.Funding.Adequacy$ –A continuous measure indicating
        whether school funding meets the needs of the district.
        Inadequate funding may correlate with higher juvenile
        delinquency rates.

        -   **Source:** School Finance Indicators Database, 2021

    -   $School.Segregation.Index$ –A measure of racial or ethnic
        segregation within schools, potentially influencing both
        educational outcomes and juvenile arrest rates.

        -   **Source**: Education Statistics Databases, 2021

    -   $Median.Prop.Tax$ –Median residential property taxes, serving as
        a proxy for socioeconomic conditions and wealth distribution
        within a county.

        -   **Source**: IPUMS, 2018-2022.

@@

-   **Outcome (Dependent) Variable** $School.Funding.Cat$

    -   Categorical variable indicating school funding adequacy. Values:
        "Adequate" or "Inadequate" (derived from
        `School.Funding.Adequacy`).
    -   **Source:** Derived from the School Finance Indicators
        Database, 2021. The adequacy metric reflects the average dollar
        gap between actual and required per-pupil spending needed to
        achieve U.S. average test scores.

-   **Primary Independent Variable:**

    -   $Residential.Segregation.Index$: A continuous variable measuring
        the level of residential segregation between Black and white
        residents in a county. Higher values indicate greater
        segregation.

        -   **Source**: American Community Survey 5-year estimates,
            2018-2022.

    -   $Median.Prop.Tax$ :A continuous variable representing the median
        property tax value in each county. It serves as a proxy for
        socioeconomic resources, as higher property taxes often
        correlate with better-funded schools.

        -   **Source**: Data from IPUMS, 2018-2022.

    -   $Median.Household.Income$ : Median income for households in the
        county (proxy needed if missing; likely not in your dataset).

        -   **Source**: Small Area Income and Poverty Estimates;
            American Community Survey, 5-year estimates, 2018-2022.

-   **Interaction Term** $Residential.Segregation.Index * Region$

    -   Examines how residential segregation’s relationship with school
        funding adequacy varies across regions.
    -   Allows the model to capture region-specific dynamics in the
        influence of segregation on school funding adequacy.

-   **Control Variables**

    -   $Spending.per.Pupil$ –Average spending per student in a
        district.

    -   $School.Segregation.Index$ –A measure of racial/ethnic
        segregation within schools.

**Load the datasets and prepare final dataset\`**:

@@

**Provide a \`glimpse():**
